Postural developments modulate children's visual access to social information

### 2020-04-29 writeup-paper/draft_apa_march2020.Rmd
updated xs-face manuscript

abstract: |
  The ability to process social information is a critical component of children's early linguistic and cognitive development.
  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver. To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
  We found that infants' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field. 
  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development.

@@

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)

@@

@@

```{r posture, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
```

@@

```{r indivPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}

@@

```{r detByAge, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}

@@

```{r detByPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\\% bootstrapped confidence intervals"}

@@

```{r detByNaming, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
```

@@

```{r franchak, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals."}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot, nrow=1)
```

>>>>

abstract: |
  The ability to process social information is a critical component of children's early linguistic and cognitive development.
  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver. To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
  We found that infants\\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field. 
  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\\' visual and linguistic experience as they change over development.

@@

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
library(png)
library(magick)

@@

### Head-mounted camera
```{r headcam, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}

img <- png::readPNG("images/viewangle_cut.png")
grid::grid.raster(img)
```

We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband. Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2]. A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, above).

[^2]: Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.

Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988]. As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).

@@

```{r posture, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
```

@@

```{r indivPosOrient,  fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}

@@

```{r detByAge, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}

@@

```{r detByPosOrient, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\\% bootstrapped confidence intervals"}

@@

```{r detByNaming, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
```

@@

```{r franchak, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals."}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot, nrow=1)
```

### 2020-04-29 writeup-paper/draft_apa_march2020.tex
updated xs-face manuscript

\documentclass[english,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Postural developments modulate children's visual access to social information},
            pdfkeywords={Postural developments modulate children's visual access to social
information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
\usepackage{csquotes}

@@

\abstract{
The ability to process social information is a critical component of
children's early linguistic and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically changing the way they see the world around them. How do
these postural and locomotor developments affect children's access to
the social information relevant for word-learning? We explored this
question by using head-mounted cameras to record the egocentric visual
perspective of 36 infants' (at 8, 12, and 16 months of age) during a
naturalistic play session with their caregiver. To estimate the
proportion of faces and hands in the infant view, we used a computer
vision algorithm to detect the presence of faces and hands in the entire
dataset. We found that infants' posture and orientation to their
caregiver changed dramatically across this age range and modulated their
access to this social information; infants who were sitting or standing
with their caregiver at a close distance tended to have the most
faces/hands in their visual field. We also applied our automated
analysis to the egocentric video data from a recent paper documenting
how posture changes one-year-olds visual access to faces (Franchak et
al., 2017), finding convergence across both methodologies and dataset
and confirming previous work that motoric developments play a
significant role in the emergence of children's linguistic and social
capacities. We suggest that the combined use of head-mounted cameras and
the application of new computer vision techniques is a promising avenue
for understanding the statistics of infants' visual and linguistic
experience as they change over development.
}

@@

@@

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/posture-1}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure*}

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/indivPosOrient-1}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure*}

@@

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/detByAge-1}

\caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure*}

\begin{figure*}[!h]

@@

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure*}

\begin{figure*}[!h]

@@

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\% bootstrapped confidence intervals.}\label{fig:detByNaming}
\end{figure*}

@@

\begin{figure*}[!h]

@@

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure*}

@@

@@

>>>>

\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Postural developments modulate children's visual access to social information},
            pdfkeywords={Postural developments modulate children's visual access to social
information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}

@@

\abstract{
The ability to process social information is a critical component of
children's early linguistic and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically changing the way they see the world around them. How do
these postural and locomotor developments affect children's access to
the social information relevant for word-learning? We explored this
question by using head-mounted cameras to record the egocentric visual
perspective of 36 infants' (at 8, 12, and 16 months of age) during a
naturalistic play session with their caregiver. To estimate the
proportion of faces and hands in the infant view, we used a computer
vision algorithm to detect the presence of faces and hands in the entire
dataset. We found that infants\textbackslash{}`posture and orientation
to their caregiver changed dramatically across this age range and
modulated their access to this social information; infants who were
sitting or standing with their caregiver at a close distance tended to
have the most faces/hands in their visual field. We also applied our
automated analysis to the egocentric video data from a recent paper
documenting how posture changes one-year-olds visual access to faces
(Franchak et al., 2017), finding convergence across both methodologies
and dataset and confirming previous work that motoric developments play
a significant role in the emergence of children's linguistic and social
capacities. We suggest that the combined use of head-mounted cameras and
the application of new computer vision techniques is a promising avenue
for understanding the statistics of infants\textbackslash{}' visual and
linguistic experience as they change over development.
}

@@

\subsubsection{Head-mounted camera}\label{head-mounted-camera}

\begin{figure}[htbp]
\centering
\includegraphics{figs/headcam-1.pdf}
\caption{\label{fig:headcam}Vertical field of view for two different headcam
configurations (we used the lower in our current study).}
\end{figure}

We used a small, head-mounted camera (\enquote{headcam}) that was
constructed from a MD80 model camera attached to a soft elastic
headband. Videos captured by the headcam were 720x480 pixels with 25
frames per second.\footnote{Detailed instructions for creating this
  headcam can be found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, above).

Even with the fish-eye lens, the vertical field of view of the camera is
still considerably reduced compared to the child's field of view, which
spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
1988; Mayer, Fulton, \& Cummings, 1988). As we were primarily interested
in the presence of faces in the child's field of view, we chose to
orient the camera upwards to capture the entirety of the child's upper
visual field where the child is likely to see adult faces, understanding
that this decision limited our ability to detect hands (especially those
of the child, which are typically found at the bottom of the visual
field).

@@

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/posture-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure}

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/indivPosOrient-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/detByAge-1} \caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure}

\begin{figure}[H]

@@

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure}

\begin{figure}[H]

@@

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\% bootstrapped confidence intervals.}\label{fig:detByNaming}
\end{figure}

@@

\begin{figure}[H]

@@

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure}

@@

\hypertarget{ref-cummings1988}{}
Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
(1988). Visual fields of young children. \emph{Behavioural and Brain
Research}, \emph{29}(1), 7--16.

@@

\hypertarget{ref-mayer1988}{}
Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
assessed with a new perimetric technique. \emph{Investigative
Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.
